---
title: "Untitled"
author: "Ashirwad Barnwal"
date: '2022-03-25'
output: html_document
---

# Getting setup
Define setup chunk:

```{r setup}
# Elegant handling of namespace conflicts
library(conflicted)

# Change function argument default
library(default)

# Path & data manipulation
library(here)
library(fs)
library(tidyverse)
conflict_prefer("filter", "dplyr")

# Pretty tables
library(sjlabelled)
library(arsenal)
library(crosstable)

## Set options

# Arsenal options
tableby_control <- tableby.control(test = FALSE, digits = 2L)
default(tableby) <- list(control = tableby_control)

## Set output paths

# Tables
tab_out_path <- here("output", "tables")
```

# Data preparation

## Import data
Import IST data:

```{r ist-raw, message = FALSE, warning = FALSE}
ist_raw <- read_csv(here("data", "IST_corrected.csv"))
glimpse(ist_raw)
```

## Clean data
Clean IST data:

```{r ist}
ist <- ist_raw %>%
  select(RDELAY:STYPE, RXASP, RXHEP, OCCODE) %>%
  mutate(
    RXHEP = fct_collapse(RXHEP, "M" = c("M", "H")),
    treatment = case_when(
      RXASP == "N" & RXHEP == "N" ~ "no_asp_no_hep",
      RXASP == "N" & RXHEP == "L" ~ "no_asp_low_hep",
      RXASP == "N" & RXHEP == "M" ~ "no_asp_med_hep",
      RXASP == "Y" & RXHEP == "N" ~ "yes_asp_no_hep",
      RXASP == "Y" & RXHEP == "L" ~ "yes_asp_low_hep",
      RXASP == "Y" & RXHEP == "M" ~ "yes_asp_med_hep"
    ),
    treatment = fct_relevel(treatment, "no_asp_no_hep") %>%
      fct_relevel("yes_asp_no_hep", after = 3),
    OCCODE = recode(
      OCCODE, `1` = "dead", `2` = "dependent", `3` = "not_recovered", 
      `4` = "recovered", .default = "missing"
    ),
    dead_or_dep = case_when(
      OCCODE %in% c("dead", "dependent") ~ "yes",
      OCCODE %in% c("not_recovered", "recovered") ~ "no",
      OCCODE == "missing" ~ "missing"
    ) %>% 
      fct_relevel("no")
  ) %>%
  select(-RXASP, -RXHEP, -OCCODE) %>%
  mutate(across(where(is.character), ~ as.factor(.x))) %>%
  filter(
    !is.na(RATRIAL), # discard pilot phase data due to incompleteness
    dead_or_dep != "missing"
  ) 
glimpse(ist)
```

# Data summary

## Variable labels
Create a tibble of variable labels:

```{r ist-var-labels}
ist_var_labels <- tibble(
  RDELAY = "Delay between stroke and randomization (hrs.)",
  RCONSC = "Conscious state at randomization",
  SEX = "Sex",
  AGE = "Age (yrs.)",
  RSLEEP = "Symptoms noted on waking",
  RATRIAL = "Atrial fibrillation",
  RCT = "CT scan before randomization",
  RVISINF = "Infarct visible on CT scan",
  RHEP24 = "Heparin within 24 hours prior to randomization",
  RASP3 = "Aspirin within 3 days prior to randomization",
  RSBP = "Systolic blood pressure at randomization (mmHg)",
  RDEF1 = "Face deficit",
  RDEF2 = "Arm/hand deficit",
  RDEF3 = "Leg/foot deficit",
  RDEF4 = "Dysphasia",
  RDEF5 = "Hemianopia",
  RDEF6 = "Visuospatial disorder",
  RDEF7 = "Brainstem/cerebellar signs",
  RDEF8 = "Other deficit",
  STYPE = "Stroke subtype",
  treatment = "Treatment",
  dead_or_dep = "Dead or dependent at 6 months follow-up?"
) %>%
  pivot_longer(everything(), names_to = "var", values_to = "label")
glimpse(ist_var_labels)
```

## Summary stats table
Generate summary stats tables of predictor variables:

```{r ist-deficit-tab, results = "asis"}
ist_deficit_tab <- ist %>%
  select(dead_or_dep, contains("DEF")) %>%
  var_labels(!!!deframe(filter(ist_var_labels, str_detect(var, "DEF")))) %>%
  mutate(
    dead_or_dep = recode(
      dead_or_dep, "no" = "Not dead/dependent", "yes" = "Dead/dependent"
    )
  ) %>%
  mutate(
    across(
      contains("DEF"), 
      ~ recode(.x, "N" = "No", "Y" = "Yes", "C" = "Can't access")
    )
  ) %>%
  tableby(dead_or_dep ~ ., data = .)
summary(ist_deficit_tab)
```

Health variables:

```{r ist-health-tab, results = "asis"}
ist_health_tab <- ist %>%
  select(dead_or_dep, RCONSC, RSLEEP, RATRIAL, RCT, RVISINF, RSBP) %>%
  var_labels(
    !!!deframe(
      filter(
        ist_var_labels, 
        var %in% c("RCONSC", "RSLEEP", "RATRIAL", "RCT", "RVISINF", "RSBP")
      )
    )
  ) %>%
  mutate(
    dead_or_dep = recode(
      dead_or_dep, "no" = "Not dead/dependent", "yes" = "Dead/dependent"
    ),
    RCONSC = recode(
      RCONSC, "F" = "Fully alert", "D" = "Drowsy", "U" = "Unconscious"
    )
  ) %>%
  mutate(across(RSLEEP:RVISINF, ~ recode(.x, "N" = "No", "Y" = "Yes"))) %>%
  tableby(dead_or_dep ~ ., data = .)
print(summary(ist_health_tab), width = 20)
```

Other variables:

```{r ist-others-tab, results = "asis"}
ist_others_tab <- ist %>%
  select(dead_or_dep, RDELAY, SEX, AGE, RHEP24, RASP3) %>%
  var_labels(
    !!!deframe(
      filter(
        ist_var_labels, var %in% c("RDELAY", "SEX", "AGE", "RHEP24", "RASP3")
      )
    )
  ) %>%
  mutate(
    dead_or_dep = recode(
      dead_or_dep, "no" = "Not dead/dependent", "yes" = "Dead/dependent"
    )
  ) %>%
  mutate(across(RHEP24:RASP3, ~ recode(.x, "N" = "No", "Y" = "Yes"))) %>%
  tableby(dead_or_dep ~ ., data = .)
print(summary(ist_others_tab), width = 20)
```

```{r export-ist-descr-tabs, include = FALSE}
tibble(
  descr_tab = list(ist_deficit_tab, ist_health_tab, ist_others_tab),
  name = list("patient-deficit", "patient-health", "patient-other")
) %>%
  pwalk(
    ~ write2word(
      .x, 
      path(tab_out_path, paste(.y, "vars-descr-stats-table.docx", sep = "-"))
    )
  )
```

## Crosstables
Create a cross table of the two treatments:

```{r ist-trt-crosstab}
ist_trt_crosstab <- ist_raw %>%
  mutate(
    RXASP = recode(RXASP, "N" = "No", "Y" = "Yes"),
    RXHEP = fct_collapse(RXHEP, "M" = c("M", "H")),
    RXHEP = fct_relevel(RXHEP, "N", "L", "M") %>%
      fct_recode("No" = "N", "Low dose" = "L", "Medium dose" = "M")
  ) %>%
  filter(!is.na(RATRIAL)) %>%
  var_labels(
    RXASP = "Trial aspirin allocated",
    RXHEP = "Trial heparin allocated"
  ) %>%
  crosstable(
    RXASP, 
    by = RXHEP, 
    total = "both", 
    percent_pattern = "{n} ({p_row}/{p_col})",
    percent_digits = 0
  ) %>%
  as_flextable()
ist_trt_crosstab
```

Create a cross table of treatment by outcome:

```{r ist-trt-by-outcome-crosstab}
ist_trt_by_outcome_crosstab <- ist %>%
  mutate(
    treatment = fct_recode(
      treatment,
      "Asp (No), Hep (No)" = "no_asp_no_hep",
      "Asp (No), Hep (Low)" = "no_asp_low_hep",
      "Asp (No), Hep (Medium)" = "no_asp_med_hep",
      "Asp (Yes), Hep (No)" = "yes_asp_no_hep",
      "Asp (Yes), Hep (Low)" = "yes_asp_low_hep",
      "Asp (Yes), Hep (Medium)" = "yes_asp_med_hep",
    ),
    dead_or_dep = fct_drop(dead_or_dep, "missing") %>%
      fct_recode("No" = "no", "Yes" = "yes")
  ) %>%
  var_labels(
    treatment = "Treatment",
    dead_or_dep = "Dead or dependent"
  ) %>%
  crosstable(
    treatment, 
    by = dead_or_dep, 
    total = "both", 
    percent_pattern = "{n} ({p_row}/{p_col})",
    percent_digits = 1
  ) %>%
  as_flextable()
ist_trt_by_outcome_crosstab
```

```{r export-ist-crosstabs, include = FALSE}
tibble(
  crosstab = list(ist_trt_crosstab, ist_trt_by_outcome_crosstab),
  name = list(
    "aspirin-by-heparin-crosstab.docx", "treatment-by-outcome-crosstab.docx"
  )
) %>%
  pwalk(
    ~ flextable::save_as_docx(.x, path = path(tab_out_path, .y))
  )
```

# Create a geographic map of IST patients

## World map
Create data for mapping:

```{r patient-map-data}
world_map <- ne_countries(returnclass = "sf")

patient_map_data <- world_map %>%
  select(name_en) %>%
  filter(name_en != "Antarctica") %>%
  mutate(name_en = recode(name_en, "United States of America" = "USA")) %>%
  left_join(patient_count_by_country, by = c("name_en" = "Country"))
glimpse(patient_map_data)
```

## Patient map
Create a map showing the geographic distribution of IST patients:

```{r patient-map}
patient_map <- patient_map_data %>%
  ggplot(aes(fill = n)) + 
  geom_sf() + 
  labs(
    subtitle = "19,435 patients from 467 hospitals in 36 countries",
    fill = NULL
  ) +
  scale_fill_viridis_c(na.value = "white") + 
  theme_bw() + 
  theme(legend.position = c(0.12, 0.15), legend.direction = "horizontal")
patient_map
```

